### 2018-09-28 paper.tex
Wip

\lstset{
  aboveskip=12pt,
  basicstyle=\footnotesize\ttfamily,
  breaklines=true,
  captionpos=b,
  frame=tb,
  framexbottommargin=5pt,
  framextopmargin=5pt,
}

@@

\setlist[description]{
  itemsep=4pt,
  labelindent=1em,
  leftmargin=1em,
  topsep=4pt,
}

\setlist[enumerate]{
  itemsep=4pt,
  topsep=4pt,
}

\setlist[itemize]{
  itemsep=4pt,
  topsep=4pt,
  labelindent=1em,
  leftmargin=1em,
}

@@

  We present Guild AI, an open source toolkit that facilitates model
  reuse for application development by applying traditional software
  packaging constructs to the domain of machine learning. Package
  management strategies are central to successful software
  ecosystems. Examples include \href{https://www.npmjs.com/}{npm} for
  JavaScript, \href{https://wiki.debian.org/Apt}{APT} for Debian, and
  \href{https://pip.pypa.io}{pip} for Python. Such tools excel at
  creating traditional software packages. Machine learning
  applications however present unique requirements that call for tool
  specialization. Guild AI combines the proven effectiveness of
  package managers with novel features to enable machine learning
  model reuse for application development across a variety of use
  cases.

@@

By running this command, a user is using a \emph{package manager}---in
this case the program \verb|pip|---to install a package named
\verb|tensorflow|. With this innocuous command, the user implicitly expects
a number of things:

@@

Package managers promote healthy software ecosystems by facilitating
software \emph{publishing}. Software authors use packaging tools to
create the installable software units---\emph{packages}---that are so
easily consumed by users. By publishing packaged software, developers
signify that their software will work as advertised if installed using
the applicable tools.

Software ecosystems that are enabled by package managers include
Python, JavaScript, Ruby, Go, and R. The GNU/Linux operating system
itself is an ecosystem of package management ecosystems including
packaging schemes from Debian, RedHat, Gentoo, and Arch
distributions. Package managers are established cornerstones of
software reuse across languages and operating systems.

@@

While models may be represented in part by traditional installed
software---e.g. an imperative TensorFlow program that defines a model
graph structure when executed---models require more to be construed as
reusable code: \emph{models must support operations on novel data}. If
they cannot be applied to novel data---either for training or for
inference---they cannot be considered reusable software.

@@

Guild AI provides various means of patching model source to support
packagers in this role. Section~\ref{sec:patching} describes patching
methods.

@@

\subsection{Manage Run Artifacts}

@@

Section~\ref{sec:runs} describes Guild AI run management.

@@

The discussion of using models to this point has not mentioned
\emph{environments} as environment are optional for using Guild
AI. However, environments are useful for isolating installed packages
and runs on a system.

@@

Environments are useful isolating project level work. It's not uncommon
to have separate environments for each project.

@@

\begin{itemize}
\item Run metadata such as the associated model and operation, status,
  start and stop times, operating system command and environment, and
  resolved dependencies
\item Output generated by the run---specifically the operating system
  process standard output and standard error
\item Files used by the operation (resolved resource sources)
\item Files generated by the operation
\end{itemize}

@@

Runs are started using the \verb|guild run| command and referencing a
model operation. Operations may be run with additional flag values
provided on the command line in the format \verb|NAME=VALUE|.

@@

As discussed above, published models---deployed from installed
packages or as projects---can be used by running model
\emph{operations}. Operations generate \emph{artifacts} that can be
used in subsequent operations or as final outputs in an application
development workflow.

In this section we discuss how such models are created using Guild AI.

@@

In many cases, this is simply a matter of selecting a model name and
creating a \verb|train| operation. Listing~\ref{lst:pet-stub}
illustrates a straightforward pet classifier.

@@

With even a simple stub, a developer can formalize the model and what
it does. This stage clarifies the model's user interface before the
functionality is implemented.

With a basic skeleton in place, the developer can implement model
operations. In the case of the pet classifier
(Listing~\ref{lst:pet-stub}) this means implementing training support
in a file named \verb|train.py| that resides in the project directory.

Operations can be developed and tested early on using Guild AI. Each
time \verb|guild run| is used to test an operation, Guild AI generates
a distinct, trackable run. Runs may be compared over time to help
answer questions or resolve issues.

@@

The first is the use the \verb|--run-dir| option
to the \verb|guild run| command. This option may be used to specify an
alternative directory for the run. This is helpful when the developer
doesn't want to generate new runs for early stage operation
development.

The second mechanism is the use of the \verb|--stage| option to the
\verb|guild run| command. This option may be used to stage a run in an
alternative directory without running the operation. This is helpful
when the developer wants to study the run directory layout or to run
low level commands or scripts during the development process.

The developer can continue to evolve the model operations as needed to
satisfy user requirements. This may include new operations or
additional operation flags, which give users more flexibility when
running an operation. The developer may similarly add new models to
the Guild file.

@@

In many cases, a developer will already have working model code she
wants to incorporate into a formalized model definition. As with the
role of \emph{packager} (Section~\ref{sec:package-maintainer}), a
developer may even be required to work with upstream sources she
cannot easily change. Guild AI provides a number of features that
support the adaptation of existing model code into usable models.

The operation \verb|main| attribute specifies a Python main module to
run. In many cases, a developer can simply reference an existing
Python script to incorporate it into an operation. In cases where a
script doesn't already exist---for example, the model code is in a
Jupyter Notebook and not readily usable---the developer should create
a new script that contains the applicable model code and use it for an
operation.

Operation \emph{flags} specified in the form \verb|NAME=VALUE| as
arguments to the \verb|guild run| command are passed to the Python
main module as command line arguments in the form
\verb|--NAME VALUE|. For example, assuming an operation \verb|train|
with a main module \verb|train|, the command
\verb|guild run train epochs=100| would be translated to the Python
command \verb|python -m train --epochs 100|. In this way developers
can parameterize scripts via operation flags.\footnote{It is standard
  practice to use Python's argparse module to support named command
  line options following the getopt convention (ref) in a Python
  script.}

If a flag name doesn't correspond to the option name expected
(e.g. the train script expected the option \verb|num_epochs|) the
developer can specify an alternative argument name for a flag using
the \verb|arg-name| attribute.

@@

Guild files may contain \verb|config| section that contain arbitrary
configuration, which can be inherited by
models. Listing~\ref{lst:inheritance} defines two models, each
inheriting the properties of shared config. The shared config
\verb|model-base| defines the common model operations and uses
\emph{parameters} to define a model name and default learning rate.

@@

\subsubsection{Configuration Includes}

@@

@@

The final method of model definition reuse described is that of
\emph{Python main module reuse}. Main modules---i.e. the modules
referenced in an operation \verb|main| specification---may reside in
any installed Python package. To reuse the main module, a developer
need only include the package reference in the main specification.

@@

Listing~\ref{lst:deploy} illustrates how such a facility could be
easily incorporated into a model to enable deployments. In the
example, the model is defined in a package that requires the
\verb|server-utils| package, which will be automatically installed by
the command \verb|guild install classifiers|.

@@

Tests can be run using the \verb|guild test| command.

Tests may be run on an ad hoc basis or as a continuous integration
release process.

\subsection{Packaging}

Our final point of discussion on the topic of model development is
\emph{packaging}. Packaging consists of generating a Python wheel
distribution that contains the Guild file and all supporting project
code. The generated wheel distribution may be distributed internally
within an organization or published to PyPI.

Packages that are published to PyPI can be installed using the command
\verb|guild install| or using pip.

Packages are convenient for installing

\section{Implementation}

Guild AI is written in the Python programming language. It makes use
of a number of excellent open source libraries, many of which are
listed below.

@@

Guild AI makes use of the Click Python library (ref) to implement its
core command line processing support.

@@

Guild View is implemented using the VueJS library (ref) for the front
end and Werkzeug (ref) for the HTTP API back end.

>>>>

\lstset{
  aboveskip=12pt,
  belowskip=0pt,
  basicstyle=\fontsize{8.5pt}{8.5pt}\ttfamily,
  breaklines=true,
  captionpos=b,
  frame=tb,
  framexbottommargin=5pt,
  framextopmargin=5pt,
}

@@

\setlist[description]{
  itemsep=0pt,
  topsep=0pt,
  labelindent=1em,
  leftmargin=1em,
}

\setlist[enumerate]{
  itemsep=0pt,
  topsep=0pt,
}

\setlist[itemize]{
  itemsep=0pt,
  topsep=0pt,
  labelindent=1em,
  leftmargin=1em,
}

@@

  We present Guild AI, an open source toolkit that facilitates model
  reuse for application development by applying traditional software
  packaging constructs to the domain of machine learning. Package
  management strategies are central to successful software
  ecosystems. Examples include npm for JavaScript \cite{npm}, APT for
  Debian \cite{apt}, and pip for Python \cite{pip}. Such tools excel
  at creating traditional software packages. Machine learning
  applications however present unique requirements that call for tool
  specialization. Guild AI combines the proven effectiveness of
  package managers with novel features to enable machine learning
  model reuse for application development across a variety of use
  cases.

@@

By running this command, a user is using a \emph{package manager}---in
this case the program \texttt|pip|---to install a package named
\verb|tensorflow|. With this innocuous command, the user implicitly expects
a number of things:

@@

\iffalse
Package managers promote healthy software ecosystems by facilitating
software \emph{publishing}. Software authors use packaging tools to
create the installable software units---\emph{packages}---that are so
easily consumed by users. By publishing packaged software, developers
signify that their software will work as advertised if installed using
the applicable tools.

Software ecosystems that are enabled by package managers include
Python, JavaScript, Ruby, Go, and R. The GNU/Linux operating system
itself is an ecosystem of package management ecosystems including
packaging schemes from Debian, RedHat, Gentoo, and Arch
distributions. Package managers are established cornerstones of
software reuse across languages and operating systems.
\fi

@@

\iffalse
While models may be represented in part by traditional installed
software---e.g. an imperative TensorFlow program that defines a model
graph structure when executed---models require more to be construed as
reusable code: \emph{models must support operations on novel data}. If
they cannot be applied to novel data---either for training or for
inference---they cannot be considered reusable software.
\fi

@@

\iffalse
Guild AI provides various means of patching model source to support
packagers in this role. Section~\ref{sec:patching} describes patching
methods.
\fi

@@

\iffalse
\subsection{Manage Run Artifacts}

@@

Section~\ref{sec:runs} describes Guild AI run management.
\fi

@@

The discussion of using models to this point has not mentioned
\emph{environments}---environment are optional for using Guild
AI. Environments are useful, however, for isolating installed packages
and runs on a system.

@@

Environments are useful isolating project level work. It is common to
have a separate environment for each project or work stream.

@@

\begin{itemize}
\item Run metadata such as the associated model and operation, status,
  start and stop times, operating system command and environment, and
  resolved dependencies
\item Output generated by the run---specifically the process standard
  output and standard error
\item Files used by the operation (resolved resource sources)
\item Files generated by the operation
\end{itemize}

@@

Runs are started using the \verb|guild run| command, referencing a
model operation. Operations may be run with additional flag values
provided in the format \verb|NAME=VALUE|.

@@

As discussed above, published models---deployed from installed
packages or as projects---are used by running
\emph{operations}. Operations generate \emph{artifacts} that can be
used in subsequent operations or used as final outputs in an
application development workflow.

In this section we discuss how models are created using Guild AI.

@@

In some cases, this may simply be a matter of selecting a model name
and creating a \verb|train| operation. Listing~\ref{lst:pet-stub}
illustrates a straightforward pet classifier.

@@

With even a simple stub, a developer can formalize the model and what
it does. This stage clarifies the model interface before the
functionality is implemented.

With a basic skeleton in place, the developer can \emph{implement}
model operations. In the case of the pet classifier in
Listing~\ref{lst:pet-stub} this means implementing train support in a
file named \verb|train.py| that resides in the project directory.

Operations can be developed and tested early using Guild AI. Each time
\verb|guild run| is used to test an operation, Guild AI generates a
distinct, tracked run. Runs may be compared over time to help answer
questions or resolve issues.

@@

The first mechanism is the use of \verb|--run-dir| with the
\verb|guild run| command. This option specifies an alternative
directory for the run. This is helpful when the developer doesn't want
to generate new runs during the development process.

The second mechanism is the use of \verb|--stage| with the
\verb|guild run| command. This option may be used to stage a run in an
alternative directory without running the operation. This is helpful
when the developer wants to study the run directory layout or to run
low level commands or scripts during the development process.

\iffalse
The developer can continue to evolve the model operations as needed to
satisfy user requirements. This may include new operations or
additional operation flags, which give users more flexibility when
running an operation. The developer may similarly add new models to
the Guild file.
\fi

@@

In many cases, a developer will already have working model code she
wants to incorporate into a formalized model definition. As with the
role of \emph{packager} (Section~\ref{sec:package-maintainer}), a
developer may be required to work with upstream sources she cannot
easily change. Guild AI provides a number of features that support the
adaptation of existing model code into usable models.

The operation \verb|main| attribute specifies a Python main module to
run. In many cases, a developer can simply reference an existing
Python script to incorporate it into an operation. In cases where a
script doesn't exist---for example, the model code is in a Jupyter
Notebook---the developer must create a new script that contains the
applicable model code and specify it for the operation.

Operation \emph{flags} specified in the form \verb|NAME=VALUE| as
arguments to the \verb|guild run| command are passed to the Python
main module as command line arguments in the form
\verb|--NAME VALUE|. For example, assuming an operation \verb|train|
with a main module \verb|train|, the command
\verb|guild run train epochs=100| would be translated to the Python
command \verb|python -m train --epochs 100|. In this way developers
can parameterize scripts using operation flags.\footnote{It is
  standard practice to use Python's argparse module to support named
  command line options following the getopt convention (ref).}

\iffalse
If a flag name doesn't correspond to the option name expected
(e.g. the train script expected the option \verb|num_epochs|) the
developer can specify an alternative argument name for a flag using
the \verb|arg-name| attribute.
\fi

@@

Guild files may contain \verb|config| sections that that can be
inherited by models definitions. Listing~\ref{lst:inheritance}
contains two models, each inheriting the properties of shared
configuration. The shared config \verb|model-base| defines the common
model operations and uses \emph{parameters} to define a model name and
default learning rate.

@@

\iffalse
\subsubsection{Configuration Includes}

@@

\fi

@@

The final method of model definition reuse discussed is that of
\emph{Python main module reuse}. Main modules---i.e. the modules
referenced in an operation \verb|main| specification---may reside in
any installed Python package. To reuse the main module, a developer
need only include the package reference in the main specification.

@@

Listing~\ref{lst:deploy} illustrates how such a facility could be
incorporated into a model to enable deployments. In the example, the
model is defined in a package that requires the \verb|server-utils|
package, which will be automatically installed by the command
\verb|guild install classifiers|.

@@

Tests are run using the \verb|guild test| command.

\iffalse
Tests may be run on an ad hoc basis or as a continuous integration
release process.
\fi

\iffalse
\section{Implementation}

Guild AI is written in the Python programming language. It makes use
of a number of excellent open source libraries, some of which are
listed below.

@@

Guild AI makes use of the Click Python library \cite{click} to
implement its core command line processing support.

@@

Guild View is implemented using the VueJS library (ref) for the front
end and Werkzeug (ref) for the HTTP API back end.
\fi
